Beyond clines: lineages and haplotype blocks in hybrid zones

### 2016-03-07 drafts/hybrid_zone.tex
added refs

There is substantial theoretical work on genetic clines.
We study those maintained \emph{only} by selection against hybrids -- \emph{tension zones} --
as in \citet{Bazykin1969}, reviewed in \citet{Barton1985}.
Clines may also be maintained by environmental gradients \citep{Haldane1948,Fisher1950};
although these are in principle different,
\citet{Slatkin1973} pointed out (among many other things)
that while the environment may determine the location of a cline,
the fitness of the heterozygote will likely determine the shape of the cline,
and \citet{Barton1986} show that the shapes of several sorts of stable clines are similar.
There is substantial work on how clines are affected by genetic drift \citep{Slatkin1975,felsenstein1975genetic,Barton2008,Polechova2011},
and quantifying the barrier to gene flow caused by clines \citep{barton1979gene,barton1989adaptation,Barton1993,Barton2000}.
In this work, we focus on \emph{transient} patterns after the initial formation of a tension zone,
in particular describing haplotype patterns.
Although the selected cline establishes quickly, transient patterns can persist for quite some time,
and haplotypes can potentially carry substantial information about the formation of recent zones.
To do this, we take the reverse-time perspective,
understanding the temporal dynamics of the system by describing properties of the \emph{lineages} tracing back from the zone.
This is particularly suited to understanding transient patterns,
although because of the well-known problems with spatial coalescence \citep{felsenstein1978pain,barton2002spatial},
incorporating genetic drift is not straightforward.
Common to nearly all studies is the use of the
reaction-diffusion equations that govern the deterministic, high-density limit \citep[][as in]{Nagylaki1975},
or the discrete analogues \citep{Hansen1966}.

>>>>

There is substantial theoretical work on genetic clines.
We study those maintained \emph{only} by selection against hybrids -- \emph{tension zones} --
as in \citet{Bazykin1969}, reviewed in \citet{Barton1985}.
Clines may also be maintained by environmental gradients \citep{Haldane1948,fisher1950frequencies};
although these are in principle different,
\citet{Slatkin1973} pointed out (among many other things)
that while the environment may determine the location of a cline,
the fitness of the heterozygote will likely determine the shape of the cline,
and \citet{Barton1986} show that the shapes of several sorts of stable clines are similar.
There is substantial work on how clines are affected by genetic drift \citep{Slatkin1975,felsenstein1975genetic,Barton2008,Polechova2011},
and quantifying the barrier to gene flow caused by clines \citep{Barton1979,Barton1989,Barton1993,Barton2000}.
In this work, we focus on \emph{transient} patterns after the initial formation of a tension zone,
in particular describing haplotype patterns.
Although the selected cline establishes quickly, transient patterns can persist for quite some time,
and haplotypes can potentially carry substantial information about the formation of recent zones.
To do this, we take the reverse-time perspective,
understanding the temporal dynamics of the system by describing properties of the \emph{lineages} tracing back from the zone.
This is particularly suited to understanding transient patterns,
although because of the well-known problems with spatial coalescence \citep{felsenstein1975pain,barton2002neutral},
incorporating genetic drift is not straightforward.
Common to nearly all studies is the use of the
reaction-diffusion equations that govern the deterministic, high-density limit \citep[][as in]{Nagylaki1975},
or the discrete analogues \citep{hanson1966effects}.
